---
title: "MIS_Adults_Effort"
author: "Kayley Dotson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(haven)
library(afex)
library(car)
library(stats)
library(plyr)
library(dplyr)
library(emmeans)
library(ppcor)
library(psych)
library(tidyr)
library(janitor)
library(kableExtra)
library(reshape2)
library(vtable)
library(summarytools)
library(knitr)
library(flextable)
library(officer)
library(ggpubr)
library(MASS)
library(coin)
```

```{r load data, include=FALSE}
setwd('/Users/kayleydotson/')
MIS_effort_adults <- read.csv("mis_adults_effort_072025.csv")
```

```{r}
dvs <- c("liking_of_high_thought_or_effort_giver", "liking_of_high_value_giver", "who_do_you_think_character_likes_more", "who_do_you_like_more", "better_friends", "which_gift_does_character_like_more", "which_gift_character_appreciates_more", "which_gift_you_like_more")
```

```{r normality1}
ggplot(MIS_effort_adults, aes(x = liking_of_high_thought_or_effort_giver)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve of Liking",
       x = "Liking",
       y = "Count")
```

```{r normality 2}
ggplot(MIS_effort_adults, aes(x = liking_of_high_value_giver)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve of Liking",
       x = "Liking",
       y = "Count")
```

```{r normality 3}
ggplot(MIS_effort_adults, aes(x = who_do_you_think_character_likes_more)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve",
       x = "Liking",
       y = "Count")
```

```{r normality 4}
ggplot(MIS_effort_adults, aes(x = who_do_you_like_more)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve",
       x = "Liking",
       y = "Count")
```

```{r normality 5}
ggplot(MIS_effort_adults, aes(x = better_friends)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve",
       x = "Liking",
       y = "Count")
```

```{r normality 6}
ggplot(MIS_effort_adults, aes(x = which_gift_does_character_like_more)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve",
       x = "Liking",
       y = "Count")
```

```{r normality 7}
ggplot(MIS_effort_adults, aes(x = which_gift_character_appreciates_more)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve",
       x = "Liking",
       y = "Count")
```


```{r normality 8}
ggplot(MIS_effort_adults, aes(x = which_gift_you_like_more)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "lightblue", color = "navy") +
  geom_density(color = "black", size = 1) +
  theme_minimal() +
  labs(title = "Histogram with density curve",
       x = "Liking",
       y = "Count")
```

Normal: better_friends, who_do_you_like_more, who_do_you_think_character_likes_more, liking_of_high_value_giver
Non-normal: which_gift_you_like_more, which_gift_character_appreciates_more, which_gift_does_character_like_more, liking_of_high_thought_or_effort_giver

## Exploratory Analysis. Predictors "location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift"

```{r}
high_effort_giver_model <- glm(liking_of_high_thought_or_effort_giver ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(high_effort_giver_model)
```

```{r}
high_value_giver_model <- glm(liking_of_high_value_giver ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(high_value_giver_model)
```

```{r}
like_who_more_model <- glm(who_do_you_think_character_likes_more ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(like_who_more_model)
```

```{r}
friends_model <- glm(better_friends ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(friends_model)
```

Age is interesting for friends model!

```{r}
who_you_like_model <- glm(who_do_you_like_more ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(who_you_like_model)
```

```{r}
recipient_gift_model <- glm(which_gift_does_character_like_more ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(recipient_gift_model)
```

```{r}
appreciate_model <- glm(which_gift_character_appreciates_more ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(appreciate_model)
```

```{r}
gift_you_like_model <- glm(which_gift_you_like_more ~ location_of_high_effort_or_thought + age_at_test + gender + character_gender + gift, family = gaussian(), data = MIS_effort_adults)
summary(gift_you_like_model)
```

N > 30, so we will proceed with t-tests.

Normal: better_friends, who_do_you_like_more, who_do_you_think_character_likes_more, liking_of_high_value_giver
Non-normal: which_gift_you_like_more, which_gift_character_appreciates_more, which_gift_does_character_like_more, liking_of_high_thought_or_effort_giver

```{r}
t.test(MIS_effort_adults$liking_of_high_thought_or_effort_giver)
```

```{r}
t.test(MIS_effort_adults$liking_of_high_value_giver)
```

```{r}
t.test(MIS_effort_adults$who_do_you_think_character_likes_more)
```

```{r}
t.test(MIS_effort_adults$better_friends)
```

```{r}
t.test(MIS_effort_adults$who_do_you_like_more)
```

```{r}
t.test(MIS_effort_adults$which_gift_does_character_like_more)
```

```{r}
t.test(MIS_effort_adults$which_gift_character_appreciates_more)
```

```{r}
t.test(MIS_effort_adults$which_gift_you_like_more)
```

```{r}
binom.test(sum(MIS_effort_adults$who_will_character_give_gift_to == 1), n = nrow(MIS_effort_adults), p = 0.5)
```

```{r}
binom.test(sum(MIS_effort_adults$you_gift == 1), n = nrow(MIS_effort_adults), p = 0.5)
```

Adults showed a preference for effort over value for each dependent variable. 
